---
title: "Satellite daily"
format: html
---

# Learning objectives  
The goals of this exercise are to:  
- Develop an API call for a specific field, dates, and bands  
- Deploy the API call and get in return the **daily** GNVDI imagery over the growing season  


# Introduction  
We are able to retrieve images in 2 ways:  

1. All images available within given dates  
2. A summary of the first option, like the mean data across all available images for a date.  

On the previous exercise, we downloaded the GNDVI **mean**.

While the mean across the growing season can be quite informative, it is still important to assess the daily imagery available.  

That will allow us to understand crop development speed over time, and which areas may have lagged behind.  

On this script, let's start the code to retrieve **all daily imagery** for our cotton field.  

Since we have 3 years of yield data (2017, 2019, 2020), let's retrieve imagery for those years, **starting with 2017**.  

# Setup  
```{r}
#install.packages("openeo")
library(openeo)
library(mapview)
library(tidyverse)
library(stars)
library(sf)
library(mapview)

```

Importing field boundary
```{r}
field <- read_sf("../../05-digitizing/ouput/boundary.geojson") %>%
  st_transform(crs = 6345) %>%
  st_buffer(dist = -10) # because values close to and outside the border are normally higher and thus skew color scales when plotting

field
```

```{r}
mapview(field)
```

# Connecting to the API  
```{r}
connection <- connect(host = "https://openeo.dataspace.copernicus.eu")
```
# Login credentials  
```{r}
login() #after running this, go to console and hit ENTER/RETURN on your keyboard to be prompted to browser to login.  
```

# Starting a process  
```{r}
p <- processes()
```


# Setting data retrieval parameters  
Here we need to determine our query parameters, including:  

- **Collection and data product** we want to pull from  

```{r}
data <- p$load_collection(id = "SENTINEL2_L2A", 
                          spatial_extent = field,
                          temporal_extent = list(
                            "2017-05-01", # start date
                            "2017-11-30" # end date
                          ),
                          bands = list("B03","B08", "SCL") #B03 = green, B08 = NIR, SCL contains land classification used to remove clouds
)

data
```

# Cloud masking  
```{r}
# define filter function to create mask
filter_function <- function(data, context) {
  vegetation <- p$eq(data["SCL"], 4) # vegetation is 4
  non_vegetation <- p$eq(data["SCL"], 5) # non-vegetation is 5
  # we want to mask all other values, so NOT (4 OR 5)
  return(p$not(p$or(vegetation, non_vegetation)))
}

# create mask by reducing bands with our defined formula
cloud_reduce <- p$reduce_dimension(data = data, 
                                   reducer = filter_function, 
                                   dimension = "bands")

# mask the bands data
cloud_reduced <- p$mask(data, cloud_reduce)
```

# Spectral reducer (GNDVI)  
```{r}

spectral_reduce <- p$reduce_dimension(data = cloud_reduced, 
                                      dimension = "bands",
                                      reducer = function(data,context) {
                                        b3 = data["B03"]
                                        b8 = data["B08"]
                                        
                                        return((b8-b3)/(b8+b3))
                                      })
```

# Compiling result  
```{r}
result_int = p$save_result(data = spectral_reduce,
                           format = "GTiff")

```

Before, for the mean image, we specified above  

`data = temporal_reduce`.  

Notice how now we just changed it to the step just before getting the mean, represented by the `spectral_reduce` object. 


# Downloading multiple individual images (before temporal reducer)  
We can also download all images.  

One of the differences here is that we need to create and start a job on the server side.

Let's first create the job below.
```{r}
job = create_job(graph = result_int,
                 title = "GNDVI",
                 description = "Growing season GNDVI")

```
Now, we can go ahead and start the job below.
```{r}
start_job(job = job)

```
Once the job is started, the images are being retrieved and processed according to our reducer steps, all on the server side.  
Once the images have been processed, we will be able to download them.

How to know when the images are processed and ready for download? 

The chunk below checks for status of the job.  
Run it a few times to check progress, it won't notify us.  
It may take 3-4 minutes.  

Once the job is completed, we can download the images.  
```{r}
list_jobs()
```

Once the job is completed, we can check its results. 
```{r}
list_results(job = job)
```

Once the job is completed, then we can download the images.  

The chunk below downloads the images from the server, and only works after the job has been completed on the server side.  

```{r}
download_results(job = job,
                 folder = "../output/sentinel/")
```

Let's check our output folder.  
Lot's of images (30)! 

Let's bring them in R to visualize.  

# Importing multiple images to check  

```{r}
rastlist = list.files(path = "../output/sentinel/",
                      pattern = ".tif$",
                      all.files = T,
                      full.names = T)

rastlist
```

```{r}
allimages_w = data.frame(source = rastlist) %>%
  mutate(date = str_sub(source, 27, 36)) %>%
  mutate(date = as.Date(date, "%Y-%m-%d")) %>%

  # Reading each imagery as a stars object  
  mutate(image_w = map(source,
                       ~read_stars(.x) %>%
                         st_transform(crs = 6345) %>%
                         rename(gndvi = 1))) %>%
  # Creating an image id
  mutate(image_id = 1:nrow(.)) %>%
  # Selecting variables
  dplyr::select(image_id, date, image_w)


allimages_w

allimages_w$image_w[[1]]
```

# EDA  
Let's visualize one of the images.  
```{r}
mapview(allimages_w$image_w[[5]])
```

# Wrangling  
Now let's crop them to the field boundary.  
```{r}
allimages_crop = allimages_w %>%
  mutate(crop = map(image_w,
                    ~.x %>%
                      st_crop(field)))

allimages_crop

mapview(allimages_crop$crop[[5]])

```

# Extracting pixel information  
Now, let's extract their pixel information.  

This will be helpful to assess pixel-level crop growth over time.  
```{r}
pixels = allimages_crop %>%
  mutate(pixel = map(crop,
                     ~.x %>%
                       st_as_sf(as_points = F
                                ))) %>%
  unnest(pixel) %>%
  group_by(date) %>%
  mutate(pixid = 1:length(gndvi)) %>%
  ungroup() %>%
  st_as_sf() %>%
  dplyr::select(date, pixid, gndvi, geometry)


pixels
```

# Maps by date  
```{r}
library(ggthemes)

pixels %>%
  ggplot() +
  geom_sf(aes(fill = gndvi,
              color = gndvi),
          size = 1) + #each one of the dates were overlaying each other
  facet_wrap(~date) +
  scale_fill_viridis_c(na.value ="transparent") +
  scale_color_viridis_c(na.value = "transparent") +
  theme_map() +
  theme(legend.position = "bottom")


ggsave("../output/timeseries.png",
       width = 5,
       height = 7,
       bg = "white")

```
Notice how we have many images with holes, representing clouds.  

```{r}
pixels %>%
  filter(date == "2017-08-18")

pixels %>%
  group_by(date) %>%
  tally()
```


# Crop development  
```{r}
pixels %>%
  filter(pixid %in% sample(1:3702, 10)) %>%
  #filter(pixid == 2000) %>%
  ggplot(aes(x = date, y = gndvi,
             group = pixid)) +
  geom_line() +
  geom_point() +
  scale_x_date(breaks = unique(pixels$date)) +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))

```

# Summary  
In this exercise, we:  
  - Retrieved and downloaded all Sentinel-2 imagery within the growing season  
  - Imported into R  
  - Processed and plotted for maps and crop growth assessment  


